---
title: "TB Seasonality Review Time Series Data Formatting"
output: html_notebook
---

Time series collected from literature will be extracted from google sheet and formatted here, then paste back to the result google sheet.

### Step 1. [Read data from google sheet](http://htmlpreview.github.io/?https://raw.githubusercontent.com/jennybc/googlesheets/master/vignettes/basic-usage.html), and filter out rows with "Keep" in "Excluded after reading full-text" column

```{r}
library(googlesheets)
library(dplyr)
df <- gs_title("Paper list")
papers <- df %>% gs_read(ws = 'Paper list')
```

```{r}
keep <- papers[papers$`Excluded after reading full-text` == 'Keep', ]
keep <- keep[!is.na(keep$RandomID), ]
```

### Step 2. Format "Start of study period" and "End of study period" column
```{r}
#standardize date formats
##keep[is.na(as.Date(keep$`Start of study period`)), ] <- paste("1", keep$`Start of study period`)

keep$`Start of study period` <- as.Date(keep$`Start of study period`)
keep$`End of study period` <- as.Date(keep$`End of study period`)
```

```{r}
keep$`Temporal scale`[keep$`Temporal scale` == "Weekly"] <- "week"
keep$`Temporal scale`[keep$`Temporal scale` == "Monthly"] <- "month"
keep$`Temporal scale`[keep$`Temporal scale` == "4-week"] <- "4 weeks"
```

### Step 3. Generate all time steps based on "Start of study period", "End of study period", and "Temporal scale" column, combine the time steps column with data column - quality control

```{r}
library(wktmo)

keep$match <- NA
keep$desired <- NA
keep$actual <- NA
library(dplyr)
keep$SubID <- with(keep, ave(rep(1, nrow(keep)), RandomID, FUN = seq_along))



pdf("H:\\__________Current Works________\\TB seasonality review\\Figures\\time series_actual.pdf", width = 8, height = 6)

for(i in 1:nrow(keep)){
  time.seq <- seq(keep$`Start of study period`[i], keep$`End of study period`[i], keep$`Temporal scale`[i])
  keep$desired[i] <- length(time.seq)
 
  time.series <- as.numeric(unlist(strsplit(keep$`Case counts`[i], split = "\\s+|\\t")))
  keep$actual[i] <- length(time.series)
  
  keep$match[i] <- keep$desired[i] == keep$actual[i]
  
  if(!is.na(time.series[1]))
    plot(time.series, xlab = "Index", ylab = "Case count", main = keep$RandomID[i], type = "l")
    points(time.series, pch = 16, cex = 0.5)
    
  # formatted.ts <- rbind(formatted.ts, data.frame(StudyID = keep$RandomID[i], SubID = keep$SubID[i], Date = time.seq, CaseCount = time.series))
}
dev.off()




```


### Step 4.  Paste the results back to the result google sheet
```{r}
df <- df %>% gs_ws_new(ws_title = "timepoints_check_1112", input = keep, trim = T, verbose = F)


```

### Step5. Formatted data to long table, convert to monthly data
```{r}
to.month <- function(time.seq, Casecount)
{
  library(lubridate)
  out.month <- seq(floor_date(time.seq[1], "month"), floor_date(time.seq[length(time.seq)], "month"), "month")
  out.month <- array(0, length(out.month), dimnames = list(as.character(out.month)))
  

  
  # add a fake time point at the end
  time.interval <- time.seq[2]-time.seq[1]
  time.seq.p1 <- c(time.seq, time.seq[length(time.seq)] + time.interval)
  
  data.month <- floor_date(time.seq.p1, "month")
  
  # distribute current case count to month
  for(i in 1:(length(time.seq.p1)-1))
  {
    if(data.month[i] == data.month[i+1])
    {
      out.month[as.character(data.month[i])] <- out.month[as.character(data.month[i])] + Casecount[i]
    } else
    {
      out.month[as.character(data.month[i])] <- out.month[as.character(data.month[i])] + round(Casecount[i]/as.numeric(time.interval)*as.numeric((ceiling_date(time.seq.p1[i], "month") - time.seq.p1[i])),0)
      out.month[as.character(data.month[i+1])] <- out.month[as.character(data.month[i+1])] + round(Casecount[i]/as.numeric(time.interval)*(as.numeric(time.interval) - as.numeric(ceiling_date(time.seq.p1[i], "month") - time.seq.p1[i])),0)
    }
    
    # print(out.month)
  }
  
  # scale the incomplete month
  # for the first month
  out.month[1] <- round(out.month[1]/(days_in_month(data.month[1]) - as.numeric(time.seq[1] - data.month[1]))*days_in_month(data.month[1]),0)
  # for the last month
  out.month[length(out.month)] <- round(out.month[length(out.month)]/(as.numeric(time.seq.p1[length(time.seq.p1)] - data.month[length(data.month)-1]))*days_in_month(data.month[length(data.month)-1]),0)
  
  if(data.month[length(data.month)] == data.month[length(data.month)-1])
  {
    out.month
  } else
  {
    out.month[-length(out.month)]
  }
  
}




formatted.ts <- NULL
for(i in 1:nrow(keep)){
  time.seq <- seq(keep$`Start of study period`[i], keep$`End of study period`[i], keep$`Temporal scale`[i])
  time.series <- as.numeric(unlist(strsplit(keep$`Case counts`[i], split = "\\s+|\\t")))
  
  if(keep$`Temporal scale`[i] != "month")
  {
    month.ts <- to.month(time.seq, time.series)
    time.seq <- as.Date(names(month.ts))
    time.series <- month.ts
  }
  
  formatted.ts <- rbind(formatted.ts, data.frame(StudyID = keep$RandomID[i], SubID = keep$SubID[i], Date = time.seq, CaseCount = time.series))
}




# very slow....
# df.ts <- gs_title("TB systematic review plan")
# df.ts <- df.ts %>% gs_edit_cells(ws = "Time series table", input = formatted.ts, trim = TRUE)
# df.ts <- df.ts %>% gs_ws_new(ws_title = "Time series table_1113", input = formatted.ts, trim = T, verbose = F)
write.csv(formatted.ts, "Formatted time series.csv", row.names = FALSE)
```

